---
title: "DB in the INFO column"
author: "Brian Knaus"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Subset to one chromosome}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Preface

This is a document prepared in 'R Markdown.'
Markdown is a language to help make simple *.html documents.
R Markdown extends Markdown to include R code which is executed when the document is compiled.
At the present, GitHub will compile the Markdown, but not the R.
To view the R you will need to compile this document locally.
If you are using R Studio you can use the 'Knit' button.



## Background


A user has contacted me regarding the query of the 'DB' field in the 'INFO' column of the 'fix' slot.
They have generated their VCF file using Mutect2.
When they query their data they are returned an `NA` and they feel this may be inappropriate.


```{r, eval=FALSE}
unique(extract_info_tidy(vcf2)$DB)
NA
```


According to the VCF specification the DB field should contain annotation information associated with a variant.
This information was presumedly provided when Mutect2 was called.
This information should come in the form of a 'key, value' pair.
The VCF specification provides for the inclusion of a key without the value as the user has presented.
The example meta line from the VCF specification indicates that there are no values for this data.
This is seen below with the '<NUMBER=0>' in the meta record.
This appears to be a flag for whether a variant is present in dbSNP and provides no further information.
We can treat this as such or add a value in our simulation to provide an example for if it were populated.


## Exploring the data


The example VCF data that comes with vcfR does not include tha 'DB' field.
But we can use it to explore where it might be if it were.


```{r}
library(vcfR)
data(vcfR_example)
head(vcf)
```


The `head()` function summarizes the VCF data.
The 'INFO' column is suppressed from this view because it is frequently rather long.
We can view it as follows.


```{r}
head(vcf@fix[,'INFO'])
```


The 'DB' field is not in this data.
We can validate this as follows.


```{r}
grep('DB', vcf@fix[,'INFO'])
```


We can repeat this with the 'tidy' method as well.


```{r}
unique(extract_info_tidy(vcf)$DB)
```


## Simulated empty data


We can simulate some 'DB' data for the purpose of troubleshooting.
The tidy function queries the meta slot for a definition.
We'll 'borrow' the meta information from the VCF specification.


```{r}
vcf@fix[,'INFO'] <- paste("DB;", vcf@fix[,'INFO'], sep="")
vcf@meta[30] <- '##INFO=<ID=DB,Number=0,Type=Flag,Description="dbSNP membership, build 129">'
```


Now that we have a 'DB' record in the 'meta' slot we can query for it.


```{r}
grep('DB', vcf@meta, value = TRUE)
```


And search for it in the 'INFO' column.


```{r}
head(grep('DB', vcf@fix[,'INFO']))
```


We can repeat this with the 'tidy' method as well.


```{r}
unique(extract_info_tidy(vcf)$DB)
```


This provides an example of behaviour encountered when the 'DB' key is present but no values are provided.


## Simulated data containing some data


As stated above, the 'DB' field does not appear to include a value.
We can add a value for the purpose of an example.
We can create an example containing values by populating some of our variants with annotations.


```{r}
data(vcfR_example)
myDB <- rep('DB', times=nrow(vcf))
myDB[20:29] <- 'DB=myAnnotation1'
myDB[50:59] <- 'DB=myAnnotation2'
vcf@fix[,'INFO'] <- paste(myDB, vcf@fix[,'INFO'], sep=";")
vcf@meta[30] <- '##INFO=<ID=DB,Number=0,Type=Flag,Description="dbSNP membership, build 129">'
```


```{r}
head(grep('DB', vcf@fix[,'INFO']))
```


We can repeat this with the 'tidy' method as well.


```{r}
unique(extract_info_tidy(vcf)$DB)
```



## Session 


As a developer I work almost exclusively on the development version of the software. 
This means that when I try to troubleshoot an issue raised by a user I need to remember to revert to the version on CRAN.
Otherwise I may be proposing solutions based on functionality that are not available in the CRAN version.
If you notice anything unusual here, please feel free to contact me.


```{r}
sessionInfo()
```
