---
title: "Heatmaps"
author: "Brian J. Knaus"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Visualization 2}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=7)
knitr::opts_chunk$set(fig.align = 'center')
```


Heatmaps can be effective way to visualize large quantities of data.
Within vcfR we've implemented our own flavor of heatmap with `heatmap.bp()`.
Many other flavors exist within R, some of which I've included in the 'See Also' section of the manual page `?heatmap.bp`.


Here we work through an example.
First we load the package and test data.


```{r}
library(vcfR)
data(vcfR_test)
```


We can explore the contents of the vcfR object by using the `head()` function.


```{r}
head(vcfR_test)
```


The element 'DP' describes how many times each variant has been sequenced.
Its also a good example of continuous numeric data.
We can use the `extract.gt()` function to create a matrix containing this data.


```{r}
dp <- extract.gt(vcfR_test, element='DP', as.numeric=TRUE)
class(dp)
dp
```


The object returned from `extract.gt()` is a matrix, a standard R container.
More information on this class can be found on its manual pages using `?matrix`.
We can now visualize this matrix with a heatmap.


```{r}
heatmap.bp(dp)
```


The column and row names are displayed with the marginal barplots.
By default, the rownames are their row numbers in the matrix.
This also illustates that row number one is at the bottom of the plot and subsequent rows appear above it.


Some may find this naming convention to be lacking.
We can rename the rownames to anything we feel is appropriate.
One option is to use the CHROM and POS information from the VCF data.


```{r}
rownames(dp) <- paste(vcfR_test@fix[,1], vcfR_test@fix[,2], sep=":")
heatmap.bp(dp)
```


The 'ID' column of the VCF data may also contain meaningful names.
Our example includes an 'ID' column that includes a mixture of names and missing data (NA).
We can use the `addID()` function to populate the missing 'ID' slots with the 'CHROM' and 'POS' information.



```{r}
vcfR_test <- addID(vcfR_test)
vcfR_test@fix
```


And use this to name our rows.


```{r}
rownames(dp) <- vcfR_test@fix[,'ID']
heatmap.bp(dp)
```




```{r}
sessionInfo()
```


