---
title: "snpclone objects"
author: "Brian J. Knaus"
date: "March 1, 2016"
output:
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{snpclone objects}
  %\VignetteEngine{knitr::rmarkdown}
  %\usepackage[utf8]{inputenc}

---







# The father, son and the snpclone object


```{r}
extends('snpclone')

getClassDef("snpclone")



#methods(snpclone)

#showMethods(clases='snpclone')

slotNames('snpclone')

#showMethods('snpclone')

methods(class='snpclone')

showMethods(class='snpclone')

```




gets
sets


indNames()
pop()


subsetting
[]




as.matrix()








## Slots


The snpclone object is based on, or extends, the genlight object.
This means that in order to understand the snpclone object, we need to understand the genlight object.
Detailed help can be found through the R help system.


```{r}
?genlight
?snpclone
```


### genlight


More information on objects of class genlight can be found in the 'genomics' tutorial at the [adegenet tutorials](https://github.com/thibautjombart/adegenet/wiki/Tutorials) page.




```{r}
my.genlight
```


### snpclone



```{r}
my.snpclone

hist(bitwise.dist(my.snpclone), col=3)

pthresh <- mlg.filter(my.snpclone)

cutoff_predictor(pthresh)


missingno(my.snpclone)

filter_stats(my.snpclone, plot=TRUE)


mlg.filter(my.snpclone, threshold=0.05)
mlg.filter(my.snpclone) <- 0.05

mlg.id(my.snpclone)



```






